---
title: "dplyr tutorial"
author: "Grace Skinner"
date: "19/May/2022"
output:
  pdf_document: default
  html_document: default
---

[Kaggle - Dive into dplyr website](https://www.kaggle.com/code/jessemostipak/dive-into-dplyr-tutorial-1/notebook) 

<br> 

### Intro

* For cleaning and exploring data
* Can feed manipulated data into ggplot2 package for visualization
* Use the Palmer Penguins dataset, which I have downloaded 

<br>

### Set up our environment 

Either load whole tidyverse package or the packages we're interested in 

```{r}
library(tidyverse)
```
<br>

Conflicts are due to multiple packages using the same name for their functions. E.g. the filter() function from the dplyr package masks the filter() function from the stats package. To use the filter() function in the stats package, we will have to use stats::filter() syntax. 

<br>

### Import our data 

```{r}
penguins <- read_csv("penguins_size.csv")
# The read_csv is a function in readr (package in tidyverse)
# Don't worry about the text appearing being in red
# Could equally just use the read.csv function 
penguins
```

<br>

# Check out our data

```{r}
# glimpse()
# Includes dataframe structure, variables and their data type, and a look at the first few rows of each variable
glimpse(penguins)

# head()
# Shows first few rows of dataframe
# Good for seeing data is formatted consistently/correctly
head(penguins)

# summary()
# summary statistics for the dataset
# Can get idea of spread, or how much missing data there is
summary(penguins)

# names()
# to get names of all variables in dataset
names(penguins)

```
<br>

### Exploring our data

%>% is the pipe operator, which allows us to push our data through sequential functions. 

<br>

## filter()
```{r}

# Take the penguins dataset and then filter for all penguins that live on the Torgersen island
penguins %>% 
  filter(island == "Torgersen")

penguins %>%
  filter(island == "Torgersen") %>%
  head()

```
<br>

## arrange()
* Arranges/organsises data in ascending order
* Provide a single argument to the arrange() function
```{r}

# numeric data
penguins %>%
  arrange(culmen_length_mm) %>%
  head()

# character data
penguins %>%
  arrange(species)

```

<br>

## Subset
```{r}

# creating a random subset of the penguins dataset
set.seed(406)

penguins_subset <- penguins %>%
  sample_n(12)  # another dplyr function!

penguins_subset

penguins_subset %>%
  arrange(species)

# Can also nest desc() inside arrange() to get data in descending order
# numeric data arranged in descending order
penguins_subset %>%
  arrange(desc(culmen_length_mm))

# character data arranged in descending - reverse alphabetical - order
penguins_subset %>%
  arrange(desc(species))

```


<br>

## More filtering 
* Can use filter on single or multiple conditions
```{r}

# filter with a single numeric condition
penguins_subset %>%
  filter(culmen_depth_mm > 16.2)

# filter with a single character condition
penguins_subset %>%
  filter(island == "Dream")

# filter with a single numeric condition between two values
penguins_subset %>%
  filter(between(culmen_depth_mm, 16.2, 18.1))

```

<br>

## select()
* Pick which columns/variables we want to look at
* Use it to pull a subset of variables, or rearrange order of variables 
```{r}

# selecting species, flipper_length_mm, and sex columns
penguins_subset %>%
  select(species, flipper_length_mm, sex)

# selecting all character data
penguins_subset %>%
  select(where(is.character))

# selecting all numeric data
penguins_subset %>%
  select(where(is.numeric))

# selecting all character data by using "where not numeric" data
penguins_subset %>%
  select(!where(is.numeric))

```

<br>

## mutate()
* Can create new columns/variables
* Works well with group_by()
```{r}

# converting grams to pounds
# notice how the order of our columns stays the same, and the new column, body_weight_pounds, gets placed at the far right of the dataframe. 
penguins_subset %>%
  mutate(body_weight_pounds = body_mass_g / 453.59237)

# Combining select and mutate
penguins_subset %>%
  mutate(body_weight_pounds = body_mass_g / 453.59237) %>%
  select(everything()) # everything selects all variables 

```

<br>

## summarise()
* Can either use summarise() or summarize()
* Useful to use with group_by()
```{r}

# summarising the average body mass of penguins, in grams
penguins_subset %>%
  summarise(avg_body_mass = mean(body_mass_g))

# since we're now summarising our data we can go ahead and use the full dataframe, since the printout will be reasonably-sized
penguins %>%
  summarise(avg_body_mass = mean(body_mass_g))

# This doesn't work very well due to NAs in the data!!!
# For now we're going to use na.rm = TRUE, but you could use filter() from the dplyr package or drop_na() from the tidyr package as well!

# summarising body mass on the entire penguins dataset while removing NA values from the calculation
penguins %>%
  summarise(avg_body_mass = mean(body_mass_g, na.rm = TRUE))

# now let's use the grouping function, group_by(), to look at the average body mass of penguins, in grams, by species
penguins %>%
  group_by(species) %>%
  summarise(avg_species_body_mass = mean(body_mass_g, na.rm = TRUE))

# now let's calculate the average body mass by species AND island
penguins %>%
  group_by(species, island) %>%
  summarise(avg_species_body_mass = mean(body_mass_g, na.rm = TRUE))

```

<br>



























