---
title: "tidyverse style guide tutorial"
author: "Grace Skinner"
date: "19/May/2022"
output:
  html_document: default
---

[Tidyverse Style Guide Website](https://style.tidyverse.org/)

<br>

# Intro
* Style guide provides consistency, making code easier to write because you need to make fewer decisions
* Have installed styler add-in, which allows you to restyle selected text/files/projects

<br>

# Files

<br>

### Names
* Meaningful
* End in .R
* Stick to numbers, letter, and _ (don't use special characters)
* If files should be run in a particular order, prefix them with numbers. 
    * If it seems likely you’ll have more than 10 files, left pad with zero
    * Later on, if missed some files, just rename all, rather than using 2a, 2b etc.
* File names that are all lower case are best due to differences between operating systems

<br>

### Internal structure
* Use commented lines of - and = to break up your file into easily readable chunks.
* If your script uses add-on packages, load them all at once at the very beginning of the file. This is more transparent than sprinkling library() calls throughout your code

```{r}
# Load data ---------------------------------------------------------------
```

<br>

# Syntax

<br>

### Object names
* Variable and function names should use only lowercase letters, numbers, and _. Use underscores (so called snake case) to separate words within a name.
* Generally, variable names should be nouns and function names should be verbs. Strive for names that are concise and meaningful (this is not easy!).

<br>

### Spacing
* Always put a space after a comma, never before, just like in regular English.
* Do not put spaces inside or outside parentheses for regular function calls.
* Place a space before and after () when used with if, for, or while.
* Place a space after () used for function arguments
* The embracing operator, {{ }}, should always have inner spaces to help emphasise its special behaviour
* Most infix operators (==, +, -, <-, etc.) should always be surrounded by spaces
    * There are a few exceptions, which should never be surrounded by spaces:
    * The operators with high precedence: ::, :::, $, @, [, [[, ^, unary -, unary +, and :.

<br>

### Commas 
**Good**
```{r}
# x[, 1]
```

**Bad**

```{r}
# x[,1]
# x[ ,1]
# x[ , 1]
```

<br>

### Parentheses
**Good**
```{r}
# mean(x, na.rm = TRUE)
```

**Bad**
```{r}
# mean (x, na.rm = TRUE)
# mean( x, na.rm = TRUE )
```

**Good**
```{r}
# if (debug) {
#   show(x)
# }
```

**Bad**
```{r}
# if(debug){
#   show(x)
# }
```

**Good**
```{r}
# function(x) {}
```

**Bad**
```{r}
# function (x) {}
# function(x){}
```

<br>

### Embracing operator
**Good**
```{r}
# max_by <- function(data, var, by) {
#   data %>%
#     group_by({{ by }}) %>%
#     summarise(maximum = max({{ var }}, na.rm = TRUE))
# }
```

**Bad**
```{r}
# max_by <- function(data, var, by) {
#   data %>%
#     group_by({{by}}) %>%
#     summarise(maximum = max({{var}}, na.rm = TRUE))
# }
```

<br>

### Infix operators 
**Good**
```{r}
# height <- (feet * 12) + inches
# mean(x, na.rm = TRUE)
```

**Bad**
```{r}
# height<-feet*12+inches
# mean(x, na.rm=TRUE)
```

**Good**
```{r}
# sqrt(x^2 + y^2)
# df$z
# x <- 1:10
```

**Bad**
```{r}
# sqrt(x ^ 2 + y ^ 2)
# df $ z
# x <- 1 : 10
```

<br>

### Function calls
* When you call a function, you typically omit the names of data arguments, because they are used so commonly. 
* If you override the default value of an argument, use the full name
* Avoid assignment in function calls

**Good**
```{r}
# mean(1:10, na.rm = TRUE)
```

**Bad**
```{r}
# mean(x = 1:10, , FALSE)
# mean(, TRUE, x = c(1:10, NA))
```

<br>

### Control flow
* Curly braces, {}, define the most important hierarchy of R code. To make this hierarchy easy to see:
    * { should be the last character on the line. Related code (e.g., an if clause, a function declaration, a trailing comma, …) must be on the same line as the opening brace.
    * The contents should be indented by two spaces.
    * } should be the first character on the line.
* If used, else should be on the same line as }.

**Good**
```{r}
# if (y < 0 && debug) {
#   message("y is negative")
# }
```

**Bad**
```{r}
# if (y < 0 && debug) {
# message("Y is negative")
# }
# 
# if (y == 0)
# {
#     if (x > 0) {
#       log(x)
#     } else {
#   message("x is negative or zero")
#     }
# } else { y ^ x }
```

<br>

### Long lines
* Strive to limit your code to 80 characters per line. This fits comfortably on a printed page with a reasonably sized font. If you find yourself running out of room, this is a good indication that you should encapsulate some of the work in a separate function.

<br>

### Semi-colons
* Don’t put ; at the end of a line, and don’t use ; to put multiple commands on one line.

<br>

### Assignment
* Use <-, not =, for assignment.

<br>

### Data
* Use ", not ', for quoting text. 
* The only exception is when the text already contains double quotes and no single quotes.

**Good**
```{r}
# "Text"
# 'Text with "quotes"'
# '<a href="http://style.tidyverse.org">A link</a>'
```

**Bad**
```{r}
# 'Text'
# 'Text with "double" and \'single\' quotes'
```

<br>

### Logical vectors
* Prefer TRUE and FALSE over T and F.

<br>

### Comments
* Each line of a comment should begin with the comment symbol and a single space: # 
* If you discover that you have more comments than code, consider switching to R Markdown.

<br>

# Functions

<br>

### Naming
* Strive to use verbs for function names

<br>

### Long lines
* Function-indent: place each argument on its own line, and indent to match the opening ( of function:
```{r}
# long_function_name <- function(a = "a long argument",
#                                b = "another argument",
#                                c = "another long argument") {
#   # As usual code is indented by two spaces.
# }
```

* OR Double-indent: Place each argument of its own double indented line.
```{r}
# long_function_name <- function(
#     a = "a long argument",
#     b = "another argument",
#     c = "another long argument") {
#   # As usual code is indented by two spaces.
# }
```

* Function indent is preferred when it fits

### return()
* Only use return() for early returns. 
* ?!?! Otherwise, rely on R to return the result of the last evaluated expression. ?!?!
* Return statements should always be on their own line 

**Good**
```{r}
# find_abs <- function(x) {
#   if (x > 0) {
#     return(x)
#   }
#   x * -1
# }
# 
# add_two <- function(x, y) {
#   x + y
# }
```

**Bad**
```{r}
# add_two <- function(x, y) {
#   return(x + y)
# }
```

<br>

### Comments
* Use comments to explain the “why” not the “what” or “how”. 
* Each line of a comment should begin with the comment symbol and a single space: # .
* Comments should be in sentence case, and only end with a full stop if they contain at least two sentences

<br>

# Pipes
* %>% should always have a space before it, and should usually be followed by a new line. After the first step, each line should be indented by two spaces

**Good**
```{r}
# iris %>%
#   group_by(Species) %>%
#   summarize_if(is.numeric, mean) %>%
#   ungroup() %>%
#   gather(measure, value, -Species) %>%
#   arrange(value)
```

**Bad**
```{r}
# iris %>% group_by(Species) %>% summarize_all(mean) %>%
# ungroup %>% gather(measure, value, -Species) %>%
# arrange(value)

```

* If the arguments to a function don’t all fit on one line, put each argument on its own line and indent
```{r}
# iris %>%
#   group_by(Species) %>%
#   summarise(
#     Sepal.Length = mean(Sepal.Length),
#     Sepal.Width = mean(Sepal.Width),
#     Species = n_distinct(Species)
#   )
```

* Assignment - a few alternative options but I like this one best:
```{r}
# iris_long <- iris %>%
#   gather(measure, value, -Species) %>%
#   arrange(-value)
```

<br>

# ggplot2
* '+' should always have a space before it, and should be followed by a new line. This is true even if your plot has only two layers. After the first step, each line should be indented by two spaces.
* If you are creating a ggplot off of a dplyr pipeline, there should only be one level of indentation.

**Good**
```{r}
# iris %>%
#   filter(Species == "setosa") %>%
#   ggplot(aes(x = Sepal.Width, y = Sepal.Length)) +
#   geom_point()
```

**Bad**
```{r}
# iris %>%
#   filter(Species == "setosa") %>%
#   ggplot(aes(x = Sepal.Width, y = Sepal.Length)) +
#     geom_point()
```

**Bad**
```{r}
# iris %>%
#   filter(Species == "setosa") %>%
#   ggplot(aes(x = Sepal.Width, y = Sepal.Length)) + geom_point()
```

* If the arguments to a ggplot2 layer don’t all fit on one line, put each argument on its own line and indent

**Good**
```{r}
# ggplot(aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
#   geom_point() +
#   labs(
#     x = "Sepal width, in cm",
#     y = "Sepal length, in cm",
#     title = "Sepal length vs. width of irises"
#   ) 
```

**Bad**
```{r}
# ggplot(aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
#   geom_point() +
#   labs(x = "Sepal width, in cm", y = "Sepal length, in cm", title = "Sepal length vs. width of irises") 
```

* ggplot2 allows you to do data manipulation, such as filtering or slicing, within the data argument. Avoid this, and instead do the data manipulation in a pipeline before starting plotting.

<br>

# Making packages

<br> 

### Files
* If a file contains a single function, give the file the same name as the function.
* If a file contains multiple related functions, give it a concise, but evocative name.

<br>

### Documentation 
* Use the first line of your function documentation to provide a concise title that describes the function, dataset, or class. Titles should use sentence case but not end with a full stop
* Always indent with one space after #'
* For most tags, like @param, @seealso and @return, the text should be a sentence, starting with a capital letter and ending with a full stop

### Error messages
* If the cause of the problem is clear, use “must” - clear cut causes typically involve incorrect types or lengths.
* If you cannot state what was expected, use “can’t”
* The problem statement should use sentence case and end with a full stop.
* Use simple sentences layed out in a bullet list of ℹ and ✖ elements
* Do your best to reveal the location, name, and/or content of the troublesome component
* If the list of issues might be long, make sure to truncate to only show the first few
* If the source of the error is clear and common, you may want to provide a hint as to how to fix it. If UTF-8 is available, prefix with ℹ. Hints should always end in a question mark.
```{r}
# dplyr::filter(iris, Species = "setosa")
# #> Error: Filter specifications must be named.
# #> ℹ Did you mean `Species == "setosa"`?
# 
# ggplot2::ggplot(ggplot2::aes())
# #> Error: Can't plot data with class "uneval". 
# #> ℹ Did you accidentally provide the results of aes() to the `data` argument?
```
* Good hints are difficult to write because, as above, you want to avoid steering users in the wrong direction. Generally, I avoid writing a hint unless the problem is common, and you can easily find a common pattern of incorrect usage
* Errors should be written in sentence case, and should end in a full stop
* If you can detect multiple problems, list up to five. This allows the user to fix multiple problems in a single pass without being overwhelmed by many errors that may have the same source

### News
* Each user-facing change to a package should be accompanied by a bullet in NEWS.md
* Strive to place the name of the function as close to the beginning of the bullet as possible

**Good**
```{r}
# * `ggsave()` now uses full argument names to avoid partial match warning (#2355).
```

**Bad**
```{r}
# * Fixed partial argument matches in `ggsave()` (#2355).
```
* Lines should be wrapped to 80 characters, and each bullet should end in a full stop.#
* Frame bullets positively (i.e. what now happens, not what used to happen), and use the present tense.

<br>


















